---
title: "Semester Project STA 6704"
author: "Benjamin Garcia"
date: "`r Sys.Date()`"
output: html_document
---

## Data Prep

```{r }
library(readr)
loan_data <- read_csv("loan_data.csv")
```


Set binary values
```{r }
loan_data$Employment_Status <- as.factor(ifelse(loan_data$Employment_Status == 'employed', 1, 0))
loan_data$Approval <- as.factor(ifelse(loan_data$Approval == 'Approved' , 1, 0))
```

For employment status, 1 if employed, 0 if not employed.
For approval, 1 if approved, 0 if not.

```{r }
table(loan_data$Approval)
(2*3933)/20067
```
balance classes

```{r}
set.seed(8675309)
ind <- sample(1:20067, 3933, replace = F)
```


```{r}
not_approved <- loan_data[loan_data$Approval == 0,]
not_approved <- not_approved[ind,]
approved <- loan_data[loan_data$Approval == 1,]
loan2 <- rbind(not_approved, approved)
```


scale numeric data
```{r}
loan3 <- cbind(loan2[,1],scale(loan2[,2:5]),loan2[,6:7])
```

## Visualizations and Variable Analysis

```{r}
library(ggplot2)
corrplot::corrplot(cor(loan3[,2:5]))

```

```{r}
plot(hclust(dist(t(loan3[,2:5]))))
```

```{r}
ggplot(loan2, aes(fill = Approval, y = Income)) + geom_boxplot()

ggplot(loan2, aes(fill = Approval, y = Loan_Amount)) + geom_boxplot()

ggplot(loan2, aes(fill = Approval, y = Credit_Score)) + geom_boxplot()

ggplot(loan2, aes(fill = Approval, y = DTI_Ratio)) + geom_boxplot()
```
pca

```{r}
data_pca <- princomp(loan3[,2:5])
```

```{r}
print(summary(data_pca))
print(summary(prcomp(loan3[,2:5])))
```

```{r}
plot(data_pca$scores[,1], data_pca$scores[,2], xlab = "PC1", ylab = "PC2", main = "First vs. Second PC")
```

```{r}
library(factoextra)
par(mfrow = c(1,1))
fviz_eig(data_pca, addlabels = TRUE)
```


```{r}
#fviz_pca_var(data_pca, col.var = "black")

#fviz_cos2(data_pca, choice = "var", axes = 1:2)

fviz_pca_var(data_pca, col.var = "cos2",
             gradient.cols = c("black", "orange", "green"),
             repel = TRUE)
```

```{r}
par(las = 3)
#barplot(data_pca$ $rotation[,"PC1"])
barplot(prcomp(loan3[,2:5])$rotation[,"PC1"])
barplot(prcomp(loan3[,2:5])$rotation[,"PC2"])
```
## Supervised Learning

logistic regression

```{r}
set.seed(8675309)
ind <- sample(1:7866, 7866*.7, replace = F)
train.df <- loan3[ind,]
holdout.df <- loan3[-ind,]


```


```{r}
logmod1 <- glm(formula = Approval ~., family = binomial(link = "logit"), data = train.df[,2:7])

summary(logmod1)
```
```{r}
step_mod_both <- MASS::stepAIC(
  object = logmod1,
  direction = "both"
)

step_mod_both
```
leave all variables in

```{r}
pred <- predict(step_mod_both, holdout.df[,2:7])

prob.predictions <- 1 / (1 + exp(-pred))

caret::confusionMatrix(factor(ifelse(prob.predictions > .5, 1, 0)), factor(holdout.df$Approval ))
```
```{r}
# ROC curve
library(ROCR)
predob <- prediction(ifelse(prob.predictions > .5, 1, 0), holdout.df$Approval )
perf <- performance(predob, "tpr", "fpr")
perf.df <- data.frame(tpr = perf@x.values[[1]],
                      fpr = perf@y.values[[1]])
ggplot2::ggplot(perf.df, aes(x = tpr, y = fpr))+
  geom_line()+
  geom_segment(aes(x=0, y=0, xend=1, yend=1), color = "gray", linetype = "dashed")+
  labs(x = "1-Specificity", y = "Sensitivity")

performance(predob, measure = "auc")@y.values[[1]]
```


random forest

```{r}
library(randomForest)
rand_for <- randomForest(Approval ~ ., data = train.df[,2:7])

```


```{r}
varImpPlot(rand_for)
```

```{r}
rf.pred <- predict(rand_for, holdout.df[,2:7], type = "class")
caret::confusionMatrix(as.factor(rf.pred), as.factor(holdout.df$Approval))
```

```{r}


```

```{r}
rownames(holdout.df) <- 1:nrow(holdout.df)
```



```{r}
#prob_bt2 <- predict(rand_for, holdout.df[,2:7])
#predob2 <- prediction(prob_bt2, holdout.df$Approval )
#perf2 <- performance(predob2, "tpr", "fpr")
#perf.df2 <- data.frame(tpr = perf2@x.values[[1]],
#                      fpr = perf2@y.values[[1]])
```
## Unsupervised Learning

bayesian

```{r}
library(Rgraphviz)
library(bnlearn)
model_hc <- bnlearn::hc(loan3[,2:5])
```


```{r}
graphviz.plot(model_hc)
```

```{r}
score(x = model_hc, data = loan_data[,2:5], type = 'bic-g')
```
```{r}
#graphviz.plot(model_hc)
strength_loan <- arc.strength(
x = model_hc,
data = loan3[,2:5])
strength.plot(x = model_hc, strength = strength_loan)
```
```{r}
bn_loan <- bn.fit(
x = model_hc,
data = loan3[,2:5]
)
```

```{r}
bn_loan_pred <- predict(
object = bn_loan,
data = loan3[,2:5],
node = colnames(loan3)[4])
```

```{r}
colnames(loan3)[4]
```


```{r}
mean((bn_loan_pred - loan3$Loan_Amount )^2)
```
clustering

```{r}
library(cluster)
library(factoextra)

#create plot of number of clusters vs total within sum of squares
fviz_nbclust(loan3[,2:5], kmeans, method = "wss")
```

```{r}
kmeans_loan <- kmeans(loan3[,2:5], centers = 2)

kmeans_loan
```

```{r}
kmeans_loan5 <- kmeans(loan3[,2:5], centers = 5)

kmeans_loan5
```

```{r}
loan5 <- cbind(loan3, kmeans_loan5$cluster)

table(loan5$Approval , loan5$`kmeans_loan5$cluster`)
```

```{r}
loan22 <- cbind(loan3, kmeans_loan$cluster)
table(loan22$Approval, loan22$`kmeans_loan$cluster`)
```



## Bag of Words

```{r}
loan3$Text <- stringr::str_to_lower(loan3$Text)

```


```{r}
loan3$Text <- qdapRegex::rm_twitter_url(
loan3$Text,
replacement = " ",
clean = TRUE
)

```


```{r}
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "ã",
replacement = "a"
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "š",
replacement = "s"
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "â",
replacement = "a"
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "¿",
replacement = "?"
)
```


```{r}
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = " i'm ",
replacement = " i am "
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "'re ",
replacement = " are "
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "'t ",
replacement = " not "
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "'ve ",
replacement = " have "
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "'ll ",
replacement = " will "
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = " doesn't ",
replacement = " does not "
)
```


```{r}
loan3$Text <- stringr::str_replace_all(
string = loan3$Text ,
pattern = "[:punct:]",
replacement = " "
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "[:digit:]",
replacement = " "
)
loan3$Text <- stringr::str_replace_all(
string = loan3$Text,
pattern = "\\W",
replacement = " "
) 


```



```{r}
loan3$Text <- tm::removeWords(
x = loan3$Text,
words = tm::stopwords(kind = "SMART")
)
loan3$Text <- tm::removeWords(
x = loan3$Text,
words = tm::stopwords(kind = "english")
)
loan3$Text <- tm::removeWords(
x = loan3$Text,
words = qdapDictionaries::Top200Words
)
```



```{r}
loan3$Text <- tm::removeWords(
x = loan3$Text,
words = c('loan', 'funds', 'money', 'financial', 'assistance', 'support', 'pay', 'purchase', 'buy', 'cover',
          'payment')
)
```


```{r}
loan3$Text <- trimws(stringr::str_replace_all(
string = loan3$Text,
pattern = "\\s+",
replacement = " "
))
```


```{r}
strsplit_text <- strsplit(loan3$Text," ")
dictionary_text <- sort(unique(unlist(strsplit_text)))
strsplit_text <- lapply(
X = strsplit_text,
FUN = tm::stemDocument
)
#strsplit_tweets <- lapply(
# X = strsplit_tweets,
# FUN = tm::stemCompletion,
# dictionary = dictionary_tweets
#)
strsplit_text <- lapply(
X = strsplit_text,
FUN = paste,
collapse = " "
)
loan3$text2 <- unlist(strsplit_text)
```


```{r}
Corpus_text <- tm::VCorpus(tm::VectorSource(loan3$text2))

DocumentTermMatrix_text <- tm::DocumentTermMatrix(Corpus_text)

DocumentTermMatrix_text <- tm::removeSparseTerms(
DocumentTermMatrix_text,
0.995
)

M <- as.matrix(DocumentTermMatrix_text)
dim(M)
```

```{r}
term_frequency <- data.frame(
Term = colnames(M),
Frequency = colSums(M),
stringsAsFactors = FALSE
)
term_frequency <- term_frequency[order(term_frequency$Frequency),]

wordcloud::wordcloud(
words = term_frequency$Term,
freq = term_frequency$Frequency,
max.words = 30,
random.order = FALSE,
colors = viridis::viridis(100)
)

```

word cloud for all loans

```{r}
approved_wc <- as.data.frame(loan3[loan3$Approval == 1, 8])
```


```{r}
app_Corpus_text <- tm::VCorpus(tm::VectorSource(approved_wc$`loan3[loan3$Approval == 1, 8]`))

app_DocumentTermMatrix_text <- tm::DocumentTermMatrix(app_Corpus_text)

app_DocumentTermMatrix_text <- tm::removeSparseTerms(
app_DocumentTermMatrix_text,
0.995
)

app_M <- as.matrix(app_DocumentTermMatrix_text)
dim(app_M)
```

```{r}
app_term_frequency <- data.frame(
Term = colnames(app_M),
Frequency = colSums(app_M),
stringsAsFactors = FALSE
)
app_term_frequency <- app_term_frequency[order(app_term_frequency$Frequency),]

wordcloud::wordcloud(
words = app_term_frequency$Term,
freq = app_term_frequency$Frequency,
max.words = 30,
random.order = FALSE,
colors = viridis::viridis(100)
)
```
word cloud for loans that were approved

```{r}
napproved_wc <- as.data.frame(loan3[loan3$Approval == 0, 8])
```


```{r}
napp_Corpus_text <- tm::VCorpus(tm::VectorSource(napproved_wc$`loan3[loan3$Approval == 0, 8]`))

napp_DocumentTermMatrix_text <- tm::DocumentTermMatrix(napp_Corpus_text)

napp_DocumentTermMatrix_text <- tm::removeSparseTerms(
napp_DocumentTermMatrix_text,
0.995
)

napp_M <- as.matrix(napp_DocumentTermMatrix_text)
dim(napp_M)
```
```{r}
napp_term_frequency <- data.frame(
Term = colnames(napp_M),
Frequency = colSums(napp_M),
stringsAsFactors = FALSE
)
napp_term_frequency <- napp_term_frequency[order(napp_term_frequency$Frequency),]

wordcloud::wordcloud(
words = napp_term_frequency$Term,
freq = napp_term_frequency$Frequency,
max.words = 30,
random.order = FALSE,
colors = viridis::viridis(100)
)
```

```{r}
app_term_frequency2 <- app_term_frequency[order(rownames(app_term_frequency)),]
napp_term_frequency2 <- napp_term_frequency[order(rownames(napp_term_frequency)),]
tot_tf <- cbind(app_term_frequency2,napp_term_frequency2$Frequency)
colnames(tot_tf)[2] <- c('approved_tf')
colnames(tot_tf)[3] <- c('not_approved_tf')

```


```{r}
tot_tf <- tot_tf[order(-tot_tf$approved_tf),]
```


```{r}
plotrix::pyramid.plot(
lx = tot_tf[1:20,3],
rx = tot_tf[1:20,2],
labels = rownames(tot_tf)[1:20],
top.labels = c("Not Approved","","Approved"),
lxcol = "red",
rxcol = "blue",
unit="count",
gap = 100
)



```

```{r}
#rep(tot_tf$Term, 2)
```


```{r}
tim <- tot_tf[1:20,]
total_tf1 <- data.frame(Term = tim$Term, Status = c(rep('Approved', 40)),
                       Count = tim$approved_tf)
total_tf2 <- data.frame(Term = tim$Term, Status = c(rep('Not Approved', 40)),
                       Count = tim$not_approved_tf)

total_tf3 <- rbind(total_tf1, total_tf2)
```


```{r}
library(ggplot2)
ggplot(total_tf3, aes(fill = Status, y=Count, x=reorder(Term, Count))) + 
    geom_bar(position="dodge", stat="identity") +
  coord_flip() +
  scale_fill_manual(values=c( 
                              "darkblue", 
                             "red")) 
```

